Data Exploration coral maps: IUCN and AquaMaps

### 2016-03-27 data_explore/data_explore_coralmaps.Rmd
writing the raster and just gonna make the map in QGIS

``` {r define_fxns, echo = FALSE}
#################################################################=
### Species Map Function ###
# This function takes a dataframe of loiczid per species ID, and a target species ID 
### (as single or vector).  Return a df of cells ready to be rasterized.  
### If AquaMaps, returned df includes aquamaps probability; if IUCN, returned
### df includes proportional area.
get_iucn_spp_map <- function(sid, spp_cells) {

  spp_map <- spp_cells %>%
    filter(iucn_sid %in% sid)
  if(nrow(spp_map) == 0) {
    message('no IUCN cells found')
    spp_map <- spp_map %>%
      as.data.frame() %>%
      select(iucn_sid, loiczid) %>%
      mutate(area_iucn = NA)
  } else {
    spp_map <- spp_map %>%
      group_by(iucn_sid, loiczid) %>%        
      summarize(area_iucn = max(prop_area))
    ### The group_by() and summarize() are to collapse cells with 
    ### overlapped polygons, fixing the duplicate 'by' issue.
    ### Note: this drops presence, sciname, and subpop columns
  }

  return(spp_map %>% unique() %>% as.data.frame())
}

#################################################################=
### Map vs Depth function
### uses ggplot to create a formatted map of species ranges, including
### Aquamaps, IUCN, and overlapped ranges; returns the plot object.
plot_depthmap <- function(spp) {
  map_iucn <- get_spp_map(spp$iucn_sid, iucn_spp_cells) %>%
    mutate(iucn_pres = ifelse(area_iucn > 0, 1, 0)) %>%
    select(-area_iucn, -iucn_sid) %>%
    unique()

  r_iucn_spp <- subs(loiczid_raster, 
                     map_iucn[ , c('loiczid', 'iucn_pres')], 
                     by = 'loiczid', 
                     which = 'iucn_pres', 
                     subsWithNA = TRUE)

  spp_pts <- as.data.frame(rasterToPoints(r_iucn_spp)) %>%
    mutate(presence = 'IUCN')

  depth_poly_df <- rgdal::readOGR(dsn = 'coralmaps', layer = 'bath_200m') %>%
    fortify() %>%
    mutate(x = long, y = lat)

  spp_plot <- ggplot(spp_pts, aes(x = x, y = y)) +
    ggtheme_map + 
    geom_raster(aes(fill = presence), alpha = .8) +
    scale_fill_manual(values = c('IUCN' = '#d95f02')) +
    geom_polygon(data = depth_poly_df, color = 'black', fill = 'yellow', alpha = .8) +
    borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
    scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
    scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2))
}

```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}

@@

  spp_plot <- plot_depthmap(spp)

  print(spp_plot)
  plotfile <- file.path(dir_fig, 'spp_maps', 
                        sprintf('%s_%s.png', 
                                spp_quad, 
                                tolower(str_replace(spp_quads[i], ' ', '_'))))
  ggsave(plotfile, height = 4.5, width = 8.7, units = 'cm')
  }

>>>>

```{r read_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}

@@

  map_iucn <- get_spp_map(spp$iucn_sid, iucn_spp_cells) %>%
    mutate(iucn_pres = ifelse(area_iucn > 0, 1, 0)) %>%
    select(-area_iucn, -iucn_sid) %>%
    unique()

  r_iucn_spp <- subs(loiczid_raster, 
                     map_iucn[ , c('loiczid', 'iucn_pres')], 
                     by = 'loiczid', 
                     which = 'iucn_pres', 
                     subsWithNA = TRUE)
  writeRaster(r_iucn_spp, 'coralmaps/oculina_varicosa.tif')

#   spp_pts <- as.data.frame(rasterToPoints(r_iucn_spp)) %>%
#     mutate(presence = 'IUCN')
#   
#   rast_fine <- loiczid_raster %>%
#     crop(extent(x = c(-135, -45), y = c(0, 45)))
#   res(rast_fine) <- 0.05

#   depth_poly <- rgdal::readOGR(dsn = 'coralmaps', layer = 'bath_200m')
#   depth_pts <- rasterize(depth_poly, rast_fine) %>%
#     rasterToPoints() %>%
#     as.data.frame() %>%
#     rename(bath200m = layer)
#   
#   spp_plot <- ggplot(spp_pts, aes(x = x, y = y)) +
#     ggtheme_map + 
#     theme(panel.background = element_rect(fill = 'lightblue')) +
# #    geom_raster(aes(fill = presence), alpha = 1) +
#     geom_raster(data = depth_pts, aes(x = x, y = y, fill = bath200m),
#                 alpha = .5) +
# #    scale_fill_manual(values = c('IUCN' = '#d95f02')) +
#     borders('world', color='gray40', fill='gray45', size = .1) +
#     coord_cartesian(xlim = c(-105, -60), ylim = c(7.5, 37.5)) +
#     scale_x_continuous(breaks = seq(-105, -60, by = 15)) +
#     scale_y_continuous(breaks = seq( 0,  45, by = 15)) +
#     theme(legend.position = 'none',
#           axis.text = element_blank())
#   
#   print(spp_plot)
#   plotfile <- file.path(dir_fig, 'spp_maps', 
#                         sprintf('%s_%s.png', 
#                                 spp_quad, 
#                                 tolower(str_replace(spp_quads[i], ' ', '_'))))
#   ggsave(plotfile, height = 4.5, width = 8.7, units = 'cm')
  }
